---
title: "Cheater Variable Removal"
author: "Jack"
date: "2025-04-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
require(openxlsx)
require(tidyverse)
```

```{r data function additions}
load("NumericVariablesWithoutYeoJ.Rdata")
load("NumericVariablesWithYeoJ.Rdata")
load("FactorVariablesClean.Rdata")
```

```{r dictionary, cache = TRUE}
path_dict = "Database/LCDataDictionary.xlsx"
dictionary = read.xlsx(path_dict, sheet = "LoanStats")
```

# Starting with numeric features:

```{r}
numeric_features = colnames(preprocessedTotalNumImpute)
dictionary |>
  filter(LoanStatNew %in% numeric_features) |> 
  View()
```

Numeric features to be removed:

| Feature         | Justification for Removal                                                                                                     |
|-----------------|-------------------------------------------------------------------------------------------------------------------------------|
| last_pymnt_amnt | This is the size of the most recent payment received for the loan, isn't something we would know at the offset                |
| recoveries      | post charge off gross recovery since charge off = default this is going to be a variable only available after knowing outcome |
| total_pymnt     | payments received to date for total amount funded, self explanatory                                                           |
| total_pymnt_inv | payments received to date for total amount funded by investors, self explanatory                                              |
| total_rec_int   | interest received to date, self explanatory                                                                                   |
| total_rec_prncp | principal received to date, self explanatory                                                                                  |

: Numeric Features to Be Removed

```{r}
numeric_features_to_be_removed = c("last_pymnt_amnt","recoveries","total_pymnt","total_pymnt_inv","total_rec_int","total_rec_prncp")
numeric_features_retained = numeric_features[!numeric_features %in% numeric_features_to_be_removed]
numeric_features_retained_no_yeoJ = preprocessedTotalNumImpute[,numeric_features_retained]
numeric_features_retained_yes_yeoJ = preprocessedTotalNumYeoJ[,numeric_features_retained]

save(numeric_features_retained_no_yeoJ,
     file = "numeric_features_retained_no_yeoJ.Rdata")
save(numeric_features_retained_yes_yeoJ,
     file = "numeric_features_retained_yes_yeoJ.Rdata")
```

# Factor Variables:

```{r}
factor_features = colnames(preprocessedTotalFactor)
dictionary |>
  filter(LoanStatNew %in% factor_features) |> 
  View()
```

| Feature    | Justification for Removal                                                                                                                        |
|------------|--------------------------------------------------------------------------------------------------------------------------------------------------|
| pymnt_plan | whether a payment plan was introduced, my interpretation is that this is introduced at a point in the loan where a borrower is struggling to pay |

```{r}
factor_features_to_be_removed = "pymnt_plan"
factor_features_retained_temp = factor_features[!factor_features %in% factor_features_to_be_removed]
factor_features_retained = preprocessedTotalFactor[,factor_features_retained_temp]

save(factor_features_retained,
     file = "factor_features_retained.Rdata")
```
