---
title: "Exploratory_Analysis"
format: html
editor: visual
---

```{r loadPackages}
require(tidyverse)
require(openxlsx)
source("helper_functions.R")
```

```{r read dataframe, cache = TRUE}
path_df = "Database/LoanStats3a.csv"
path_dict = "Database/LCDataDictionary.xlsx"
df = read.csv(path_df)
#sheets = getSheetNames(path_dict)
dictionary = read.xlsx(path_dict, sheet = "LoanStats")
```

```{r}
get_description("total")
```

# Looking at Purposes

```{r}
unique_purposes = df |> 
  distinct(purpose)
unique_purposes
```

```{r}
get_description("purpose")
```

The purpose column is provided by the borrower to categorize their request. There are `r nrow(unique_purposes)` unique values.

```{r}
purpose_numbers <- df |> 
  group_by(purpose) |> 
  summarize(n = n(), .groups = "drop") |> 
  arrange(desc(n))
purpose_numbers
```

```{r}
dc_rows = as.integer(purpose_numbers |> filter(purpose == "debt_consolidation") |> select(n))
total_rows = nrow(df)
```

The most common purpose is debt consolidation, making up `r 100*(dc_rows/total_rows)` % of the total rows in the data set.

```{r, cache=TRUE}
set.seed(10)
description_examples_dc = df |> 
  filter(purpose == "debt_consolidation") |> 
  slice_sample(n = 20, replace = FALSE) |> 
  select(desc)
description_examples_dc
```

After taking a look at a few of the descriptions, many mention consolidating credit card debt. The debt consolidation purpose may overlap with the credit card purpose.

```{r}
set.seed(10)
description_examples_cc = df |> 
  filter(purpose == "credit_card") |> 
  slice_sample(n = 20, replace = FALSE) |> 
  select(desc)
description_examples_cc
```

From this sample, it appears that the credit card purpose could be considered a subset of the debt consolidation purpose. Most are simply consolidating their credit card debt.

We could loo
