---
title: "Logistic Regression"
author: "Jack"
date: "2025-04-03"
output: html_document
---

# Loading Packages and Data

```{r loadingPackages}
require(dplyr)
require(caret)
require(pROC)
require(corrplot)
```

Adding testing data:

```{r}
load("Xtest.Rda")
load("Ytest.Rda")
```

Adding training data:

```{r}
load("no_smote_train.Rda")
load("Ytrain_nosmote.Rda")
```

Since logistic regression will suffer when predictors are highly correlated, we check the correlation plot for numeric variables.

```{r}
train_df = as.data.frame(no_smote_train)
test_df  = as.data.frame(testdata)
```

Getting numeric variables to correlation filter:

```{r}
train_df_numeric = train_df[,1:18]
```

# Correlation Filtering

```{r}
train_df_numeric_cor = cor(train_df_numeric)
corrplot(train_df_numeric_cor, order = "hclust",tl.cex = .7)
```

There is a large pocket of correlation, mainly relating to the size of the loan amount.

Since there is evidence of high correlation between features, I decided to go with a correlation filter.

```{r}
highCorr = findCorrelation(train_df_numeric_cor, .8, verbose = TRUE, names = TRUE)
save(highCorr,
     file = "highCorr.Rda")
```

Filtering out highly correlated filters

```{r}
train_df = train_df[,!colnames(train_df) %in% highCorr]
test_df  = test_df[,!colnames(test_df) %in% highCorr]
```

# Logistic Regression

```{r}
trControl    = trainControl(method = 'none')
outLogistic  = train(x = train_df, y = Ytrain_nosmote,
                    method = 'glm', trControl = trControl)
YhatTestProb = predict(outLogistic, testdata, type = 'prob')
```

# Calibration Check

```{r}
YtestRelevel = relevel(Ytest,  ref = 'Charged Off')
```

```{r}
calibPlot  = calibration(YtestRelevel ~ YhatTestProb$`Charged Off`, cuts = 5)
xyplot(calibPlot)
```

Note that this calibration plot is a bit misleading, our model predicts a probability of default of .8 one time. Case 8164 in Ytest, which is charged off (hence observed probability is 100).

```{r}
sum(YhatTestProb$`Charged Off` > .8)
which(YhatTestProb$`Charged Off` > .8)
Ytest[which(YhatTestProb$`Charged Off` > .8)]
```

# Classifications

```{r}
rocCurve = roc(Ytest, YhatTestProb$`Charged Off`)
plot(rocCurve, legacy.axes = TRUE)
```

```{r}
roc_logistic_data = cbind(rocCurve$thresholds,
                          rocCurve$sensitivities,rocCurve$specificities)
save(roc_logistic_data,
     file = "roc_logistic_data.Rda")
```
