---
title: "Logistic Regression"
author: "Jack"
date: "2025-04-03"
output: html_document
---

# Loading Packages and Data

```{r loadingPackages}
require(dplyr)
require(caret)
require(pROC)
require(corrplot)
```

Adding non-transformed data:

```{r}
load("NumericVariablesWithoutYeoj.Rdata")
load("FactorVariablesClean.Rdata")
```

Since logistic regression will suffer when predictors are highly correlated, we check the correlation plot for numeric variables.

# Correlation Filtering

```{r}
NumericVariablesCorr = cor(preprocessedTotalNumImpute)
corrplot(NumericVariablesCorr, order = "hclust",tl.cex = .7)
```

There is a large pocket of correlation, mainly relating to the size of the loan amount. In that pocket there is a portion that cheats its way to the status output, in particular ones tracking the total amount paid into the loan.

Since there is evidence of high correlation between features, I decided to go with a correlation filter.

```{r}
highCorr = findCorrelation(NumericVariablesCorr, .8, verbose = TRUE, names = TRUE)
```

Grabbing numeric features after highly correlated
